---
title: "Examine topic modeling results with k = 16 topics"
author: Peter Carbonetto
output: workflowr::wflow_html
---

*Add text here giving an overview of this analysis.*

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,results = "hold",
                      fig.align = "center",dpi = 120)
```

Load the packages used in the analysis.

```{r load-pkgs, message=FALSE}
library(data.table)
library(fastTopics)
library(ggplot2)
library(cowplot)
source("../code/lps_data.R")
```

Initialize the sequence of pseudorandom numbers.

```{r set-seed}
set.seed(1)
```

Load the count data.

```{r load-data}
dat <- read_lps_data("../data/raw_read_counts.csv.gz")
samples <- dat$samples
counts  <- dat$counts
```

Remove genes with very low (or no) expression.

```{r prepare-data}
j <- which(colSums(counts) > 20)
counts <- counts[,j]
```

Load the results of the topic modeling analysis.

```{r load-results}
load("../output/fit-lps-k=16.RData")
```

Plot the improvement in the solution over time.

```{r plot-progress, fig.height=2.25, fig.width=5}
p1 <- plot_progress(fit,x = "timing",y = "loglik",colors = "black",
                    add.point.every = 10,e = 1e-4) +
  guides(color = "none",fill = "none",shape = "none",
         linetype = "none",size = "none")
p2 <- plot_progress(fit,x = "timing",y = "res",colors = "black",
                    add.point.every = 10,e = 1e-4) +
  guides(color = "none",fill = "none",shape = "none",
         linetype = "none",size = "none")
plot_grid(p1,p2)
```

Visualize the structure identified in each of the tissues using a
Structure plot, in which the samples in each tissue are ordered by
time in which the sample was taken:

```{r structure-plot-1, fig.height=1.75, fig.width=8}
set.seed(1)
rows <- order(samples$timepoint)
topic_colors <- c("darkblue","dodgerblue","darkorange","forestgreen",
                  "limegreen","tomato","darkred","olivedrab","magenta",
				  "darkmagenta","sienna","royalblue","lightskyblue",
                  "gold","red","cyan")
p <- structure_plot(fit,grouping = samples$tissue,gap = 5,
                    colors = topic_colors,
                    topics = c(15,3,4,5,6,7,8,9,10,11,12,13,2,14,1,16),
                    loadings_order = rows) +
  theme(legend.key.height = unit(0.15,"cm"),
        legend.text = element_text(size = 7))
print(p)
```

There is a single topic (topic 9, blue in the plot below) that is
capturing changes in expression over time across many tissues. Two
other topics (topics 1 and 6) show similar patterns, except these
patterns are specific to two tissues (PBMC and LI).

```{r structure-plot-2, fig.height=1.75, fig.width=8}
set.seed(1)
topic_colors <- c("gold","darkorange","dodgerblue","white")
fit2 <- poisson2multinom(fit)
fit2 <- merge_topics(fit2,paste0("k",setdiff(1:16,c(1,6,9))))
p <- structure_plot(fit2,grouping = samples$tissue,gap = 5,
                    colors = topic_colors,topics = c(4,1:3),
                    loadings_order = rows)
print(p)
```

These volcano plots summarize the results of the DE analysis for
topics capturing different tissues (and topic 9, which is capturing
changes in expression at different time points):

```{r volcano-plots-1, fig.height=18, fig.width=8, warning=FALSE}
topics <- colnames(de_merged$z)
p <- vector("list",13)
names(p) <- topics
for (k in topics) {
  p[[k]] <- volcano_plot(de_merged,k = k,ymax = 500) +
    guides(fill = "none")
  volcano_plotly(de_merged,k = k,ymax = 500,
                 file = paste("volcano_plot_",k,".html",sep = ""))
}
do.call("plot_grid",c(p,list(ncol = 3,nrow = 5)))
```

These volcano plots summarize the results of the DE analysis for topics
capturing expression patterns within tissues:

```{r volcano-plots-2, fig.height=8, fig.width=8, warning=FALSE}
topics <- c("k1","k2","k5","k6","k13","k14")
p <- vector("list",6)
names(p) <- topics
for (k in topics) {
  p[[k]] <- volcano_plot(de,k = k,ymax = 300) +
    guides(fill = "none")
  volcano_plotly(de,k = k,ymax = 300,
                 file = paste("volcano_plot_",k,".html",sep = ""))
}
do.call("plot_grid",c(p,list(ncol = 3,nrow = 2)))
```

These results may also be browsed interactively:
[k1](volcano_plots/volcano_plot_k1.html), 
[k2](volcano_plots/volcano_plot_k2.html), 
[k3](volcano_plots/volcano_plot_k3.html), 
[k4](volcano_plots/volcano_plot_k4.html), 
[k5](volcano_plots/volcano_plot_k5.html), 
[k6](volcano_plots/volcano_plot_k6.html), 
[k7](volcano_plots/volcano_plot_k7.html), 
[k8](volcano_plots/volcano_plot_k8.html), 
[k9](volcano_plots/volcano_plot_k9.html), 
[k10](volcano_plots/volcano_plot_k10.html), 
[k11](volcano_plots/volcano_plot_k11.html), 
[k12](volcano_plots/volcano_plot_k12.html), 
[k13](volcano_plots/volcano_plot_k13.html), 
[k14](volcano_plots/volcano_plot_k14.html), 
[k15](volcano_plots/volcano_plot_k15.html), 
[k16](volcano_plots/volcano_plot_k16.html), 
[k1+k5](volcano_plots/volcano_plot_k1+k5.html), 
[k2+k13](volcano_plots/volcano_plot_k2+k13.html), 
[k6+k14](volcano_plots/volcano_plot_k6+k14.html).
