---
title: "Structure plots from topic model with k = 16 topics"
author: Peter Carbonetto
output: workflowr::wflow_html
---

Here we create Structure plots showing topics 1 through 16 in
detail. See [here](examine_topic_model_k16.html) for the Structure
plot showing all topics at once.

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,results = "hold",
                      fig.align = "center",dpi = 120)
```

Load the packages used in the analysis and set the seed.

```{r load-pkgs, message=FALSE}
library(data.table)
library(fastTopics)
library(ggplot2)
library(cowplot)
source("../code/lps_data.R")
set.seed(1)
```

Load the count data.

```{r load-data}
dat <- read_lps_data("../data/raw_read_counts.csv.gz")
samples <- dat$samples
counts  <- dat$counts
```

Remove genes with very low (or no) expression.

```{r prepare-data}
j <- which(colSums(counts) > 20)
counts <- counts[,j]
```

Load the results of the topic modeling analysis.

```{r load-results}
load("../output/fit-lps-k=16.RData")
fit <- poisson2multinom(fit)
```

Structure plots for topics 1--4:

```{r structure-plots-1, fig.height=5, fig.width=6.5}
rows <- order(samples$timepoint)
topic_colors <- c("darkblue","dodgerblue","darkorange","forestgreen",
                  "limegreen","tomato","darkred","olivedrab","magenta",
                  "darkmagenta","sienna","royalblue","lightskyblue",
                  "gold","red","cyan")
plots <- vector("list",16)
names(plots) <- paste0("k",1:16)
for (k in 1:16) {
  fit2 <- merge_topics(fit,paste0("k",setdiff(1:16,k)))
  colnames(fit2$L) <- c(paste0("k",k),"other")
  plots[[k]] <- structure_plot(fit2,grouping = samples$tissue,gap = 5,
                               colors = c(topic_colors[k],"white"),
                               topics = c(2,1),loadings_order = rows) +
    guides(color = "none",fill = "none") +
	labs(y = "topic prop.",title = paste("topic",k)) 
  outfile <- paste0("structure_plot_k",k,".pdf")
  ggsave(outfile,plots[[k]],height = 1.25,width = 6.5)
}
do.call(plot_grid,c(plots[1:4],list(nrow = 4,ncol = 1)))
```

Structure plots for topics 5--8:

```{r structure-plots-2, fig.height=5, fig.width=6.5}
do.call(plot_grid,c(plots[5:8],list(nrow = 4,ncol = 1)))
```

Structure plots for topics 9--12:

```{r structure-plots-3, fig.height=5, fig.width=6.5}
do.call(plot_grid,c(plots[9:12],list(nrow = 4,ncol = 1)))
```

Structure plots for topics 13--16:

```{r structure-plots-4, fig.height=5, fig.width=6.5}
do.call(plot_grid,c(plots[13:16],list(nrow = 4,ncol = 1)))
```
